Taking Dyads Seriously

### 2018-03-02 paper/netsMatter.tex
minor changes to abstract

\small{\singlespacing{
\begin{abstract}
Much of international relations scholarship concerns dyads: dyadic hypotheses and especially dyadic data. These types of data do not typically satisfy the conditional independence criterion required for most statistical approaches. Consequently, modeling approaches that fail to adequately capture the data generating process behind dyadic events will suffer from biased coefficients and poorly calibrated standard errors. We introduce a regression-based approach, the Additive and Multiplicative Effects (AME) model, that more fully accounts for the dependencies in dyadic data. First, we conduct a simulation that reveals how the model captures dependencies that emerge in relational data and demonstrate why accounting for these processes improves inference. Second, through replication, we compare the AME model to approaches used in three studies from recent international relations scholarship. For each study, we find that the modeling approach used in the original study performs notably worse at capturing the data generating process compared to the AME. Further, these conventional methods overstate the effect of key variables and underestimate the uncertainty in these effects. In sum, our work shows the consequences of failing to seriously consider dependencies inherent to dyadic data. Perhaps most importantly, by precisely modeling the data generating process underlying political phenomena we improve scholars’ ability to conduct inferential analyses.

\end{abstract}
}}

>>>>

\small{\singlespacing{
\begin{abstract}
	Much of international relations scholarship concerns dyads: dyadic hypotheses and especially dyadic data. Yet, modeling approaches that fail to properly capture the data generating process behind dyadic events will suffer from biased coefficients and poorly calibrated standard errors. We introduce a regression-based approach, the Additive and Multiplicative Effects (AME) model, that better accounts for the dependencies in dyadic data. First, we conduct a simulation to highlight how the model captures dependencies and show that accounting for these processes improves our ability to conduct inference. Second, we compare the AME model to approaches used in three studies from recent international relations scholarship. For each study, we find that the modeling approach used in the original study performs notably worse at capturing the data generating process compared to the AME. Further, conventional methods misstate the effect of key variables and the uncertainty in these effects. In sum, our work shows the consequences of failing to take the dependencies inherent to dyadic data seriously. Most importantly, by better modeling the data generating process underlying political phenomena, the AME framework improves scholars' ability to conduct inferential analyses. 
\end{abstract}
}}
